---
title: "Historical Temperatures for Three Cities"
subtitle: "Land temperature from 1760 to 2013 for Three Cities"
author: "Greig Robertson"
date: "11/23/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Source Data

The temperature data was downloaded from the following Berkeley Earth project: "Climate Change: Earth Surface Temperature Data - Exploring global temperatures since 1750".  More details are available here:

- Berkeley Earth Project - http://berkeleyearth.org/data/
- Source data downloaded from - https://www.kaggle.com/berkeleyearth/climate-change-earth-surface-temperature-data
- The data set used from this was - GlobalLandTemperaturesByMajorCity.csv
- License for use of data - https://creativecommons.org/licenses/by-nc-sa/4.0/


## Plot Details

The plot on the next slide shows land temperature from 1760 to 2013 (where available) for Berlin, Paris and Sydney.  A loess smoothing line is fitted through the markers for each city.

You can:

- Zoom in by click-drag
- toggle markers and lines by clicking on the legend.

## Historical Temperatures

```{r plotly, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse) # for filtering, summarising
library(lubridate) # for date processing
library(plotly)    # for plotting

# Load the downloaded data set
temp_data      <- read_csv("~/data/climate-change-earth-surface-temperature-data/GlobalLandTemperaturesByMajorCity.csv")

# Set up the cities that we want to plot
cities         <- c("Berlin", "Paris", "Sydney")

# Extract the temperature data for those cities
# (note that temperatures around 1750 are very low, so this starts at 1760)
city_temp      <- temp_data %>% filter(City %in% cities & dt > as.Date("1759-12-31"))

# Aggregate temperature by year and by city (to reduce the size of the plot)
city_temp_aggr <- city_temp %>% mutate(Year = year(dt))            %>%
                                filter(!is.na(AverageTemperature)) %>%
                                group_by(Year, City)               %>%
                                summarise(AverageYearTemperature = mean(AverageTemperature))

# Create a base Plotly object
# setting the X & Y axis labels and legend postion
p <- plot_ly(colors = "Set1") %>%
      layout(xaxis  = list(title = 'Year'),
             yaxis  = list(title = 'Mean Annual Temperature (Celsius)'),
             legend = list(x = 0.05, y = 0.9, bgcolor = "#F0F0F0"))

# For each city:
#   extract the temperature data per year
#   fit a loess line through the temperature data
#   Note - need to work out how to get the markers and lines the same colour
for (city in cities) {

  city_data   <- city_temp_aggr %>% filter(City == city)

  # note that changing the span increases/decreases the bumpiness of the line
  smooth_line <- fitted( loess(city_data$AverageYearTemperature ~ city_data$Year, span = 0.3) )

  # Add a marker for the temperature measurement
  p <- add_markers(p,
                   x          = city_data$Year,
                   y          = city_data$AverageYearTemperature,
                   showlegend = TRUE,
                   name       = city)

  # Add the calculated loess smooth line
  p <-  add_lines(p,
                  x          = city_data$Year,
                  y          = smooth_line,
                  showlegend = TRUE,
                  name       = paste(city, "loess"))
}

p
```
